---
title: "Correlation: Likelihood and P-value Functions"
description: "Constructing a profile likelihood function for a correlation coefficient and pairing it with the consonance (P-value) function from curve_corr()."
output:
  rmarkdown::html_vignette:
    toc: true
opengraph:
  image:
    src: "https://res.cloudinary.com/less-likely/image/upload/v1591475692/Site/lfunctions.png"
  twitter:
    card: summary
    creator: "@dailyzad"
bibliography: references.bib
link-citations: yes
csl: american-medical-association.csl
vignette: >
  %\VignetteIndexEntry{Correlation: Likelihood and P-value Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  message = TRUE,
  warning = TRUE,
  collapse = TRUE,
  comment = "#>",
  fig.height = 4.5,
  fig.width = 6
)
library(concurve)
```

Most of the other vignettes start from a fitted model and read the
likelihood off an object built by another package. Here we do the
opposite: we build a **likelihood function from scratch** for a parameter
that `concurve` does not have a dedicated interval function for --- the
Pearson correlation coefficient $\rho$ --- and then show that the
consonance (P-value) function produced by `curve_corr()` tells the same
story from the frequentist side.[@schwederConfidenceLikelihood2002;@royallStatisticalEvidenceLikelihood1997]

The point is methodological. A likelihood function, a consonance
function, a confidence distribution, and a surprisal function are
different coordinate views of the *same* underlying object.[@rafiSemanticCognitiveTools2020;@birnbaumUnifiedTheoryEstimation1961]
If you can build any one of them, you can build the rest, and for a
one-parameter target they will very nearly coincide.

# The data

We simulate two variables with a moderate population correlation so the
example is reproducible.

```{r echo=TRUE}
set.seed(1031)
n <- 30
x <- rnorm(n)
y <- 0.6 * x + sqrt(1 - 0.6^2) * rnorm(n)

(r <- cor(x, y))
```

# Constructing the likelihood function

For a bivariate normal, the log-likelihood at a fixed value of $\rho$,
with the four nuisance parameters (the two means and two standard
deviations) profiled out by maximum likelihood, is the **profile
log-likelihood** of $\rho$.[@schwederConfidenceLikelihood2002] We compute
it directly: for each candidate $\rho$ we maximize the bivariate-normal
log-likelihood over the nuisance parameters with `optim()`.

```{r echo=TRUE}
# Profile log-likelihood of rho: maximize over (mu_x, mu_y, log sd_x, log sd_y)
prof_loglik_rho <- function(rho, x, y) {
  n <- length(x)
  neg_ll <- function(par) {
    mux <- par[1]
    muy <- par[2]
    sx <- exp(par[3])
    sy <- exp(par[4])
    dx <- (x - mux) / sx
    dy <- (y - muy) / sy
    quad <- (dx^2 - 2 * rho * dx * dy + dy^2) / (1 - rho^2)
    ll <- -n * log(2 * pi) - n * log(sx) - n * log(sy) -
      0.5 * n * log(1 - rho^2) - 0.5 * sum(quad)
    -ll
  }
  start <- c(mean(x), mean(y), log(sd(x)), log(sd(y)))
  -optim(start, neg_ll, method = "BFGS")$value
}
```

Now evaluate it over a grid of $\rho$ and coerce the result into the
data frame layout that `ggcurve()` expects for likelihood functions ---
the same five columns `curve_lik()` produces: `values`, `likelihood`,
`loglikelihood`, `support`, and `deviancestat`.

```{r echo=TRUE}
grid <- seq(-0.95, 0.95, length.out = 600)
ll <- vapply(grid, prof_loglik_rho, numeric(1), x = x, y = y)

loglik_rel <- ll - max(ll) # relative log-likelihood (max = 0)
support <- exp(loglik_rel) # relative (normalized) likelihood, max = 1

lik_rho <- data.frame(
  values        = grid,
  likelihood    = exp(ll), # raw profile likelihood         -> type = "l3"
  loglikelihood = loglik_rel, # relative log-likelihood         -> type = "l2"
  support       = support, # relative likelihood (0, 1]      -> type = "l1"
  deviancestat  = -2 * loglik_rel # -2 log rel. likelihood       -> type = "d"
)
class(lik_rho) <- c("data.frame", "concurve")
head(lik_rho)
```

This is exactly the object `curve_lik()` hands back, so every downstream
`concurve` verb works on it unchanged.

## Plotting the likelihood

```{r echo=TRUE}
ggcurve(lik_rho,
  type = "l1", nullvalue = TRUE,
  xaxis = expression(rho),
  title = "Relative Likelihood Function",
  subtitle = "Profile likelihood for the correlation coefficient"
)

ggcurve(lik_rho,
  type = "l2",
  xaxis = expression(rho),
  title = "Log-Likelihood Function"
)

ggcurve(lik_rho,
  type = "d", nullvalue = 0,
  xaxis = expression(rho),
  title = "Deviance Function"
)
```

The maximum of the relative likelihood sits at the sample correlation
$r$, and the null value $\rho = 0$ receives whatever support the data
happen to give it --- not a verdict, a height.[@pooleConfidenceIntervalsExclude1987]

## Reading support intervals

A pure-likelihood interval collects every $\rho$ whose relative
likelihood exceeds a cutoff $1/k$.[@royallStatisticalEvidenceLikelihood1997]
Royall's conventional cutoffs are $1/8$ and $1/32$; the $1/6.8$ cutoff is
the likelihood interval that, for an approximately normal likelihood,
matches the usual 95% interval (since $e^{-1.96^2 / 2} \approx 1/6.8$).

```{r echo=TRUE}
support_interval <- function(df, k) {
  keep <- df$values[df$support >= 1 / k]
  c(lower = min(keep), upper = max(keep))
}

rbind(
  `1/6.8` = support_interval(lik_rho, 6.8),
  `1/8`   = support_interval(lik_rho, 8),
  `1/32`  = support_interval(lik_rho, 32)
)
```

# The companion: consonance and surprisal functions

Now the frequentist view. `curve_corr()` inverts `cor.test()` at every
level to produce the consonance (P-value) function and its surprisal
transform.[@fraserPvalueFunctionStatistical2019;@rafiSemanticCognitiveTools2020]

```{r echo=TRUE}
rho_curve <- curve_corr(
  x = x, y = y,
  alternative = "two.sided",
  method = "pearson"
)
```

```{r echo=TRUE}
ggcurve(rho_curve[[1]],
  type = "c", nullvalue = TRUE,
  xaxis = expression(rho),
  title = "Consonance (P-value) Function"
)

ggcurve(rho_curve[[1]],
  type = "s", nullvalue = TRUE,
  xaxis = expression(rho),
  title = "Surprisal (S-value) Function"
)
```

The consonance function peaks at $r$, and its horizontal slice at
$1 - \alpha = 0.95$ is the ordinary 95% interval. The surprisal function
$S(\rho) = -\log_2 p(\rho)$ rescales the same curve into **bits of
information** against each value of $\rho$.[@rafiSemanticCognitiveTools2020]

## The two views agree

The $1/6.8$ likelihood interval and the 95% consonance interval should
land in nearly the same place --- that is the whole point.

```{r echo=TRUE}
ci95 <- as.numeric(rho_curve[[1]][
  which.min(abs(rho_curve[[1]]$intrvl.level - 0.95)),
  c("lower.limit", "upper.limit")
])

rbind(
  `Likelihood 1/6.8` = support_interval(lik_rho, 6.8),
  `Consonance 95%`   = c(lower = ci95[1], upper = ci95[2])
)
```

They are close but not identical: the likelihood interval is exact for
the profiled normal model, while `cor.test()` uses Fisher's $z$
approximation. The small gap is a feature to notice, not a bug to hide
--- it is the difference between two legitimate inferential
summaries.[@schwederConfidenceLikelihood2002]

# Cite R Packages

Please remember to cite the packages that you use.

```{r}
citation("concurve")
```

# References

* * *
