---
title: "Customizing Plots"
description: "A guide on how to customize the plots obtained from concurve via ggplot2 arguments and with other packages."
output: 
  rmarkdown::html_vignette:
    toc: true
opengraph:
  image: 
    src: "https://upload.wikimedia.org/wikipedia/commons/2/29/Minard.png"
  twitter:
    card: summary
    creator: "@dailyzad"
bibliography: references.bib
link-citations: yes
csl: american-medical-association.csl
vignette: >
  %\VignetteIndexEntry{Customizing Plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  message = TRUE,
  warning = TRUE,
  collapse = TRUE,
  comment = "#>"
)
```

# A Universal Grammar of Graphics

Because `concurve` graphs functions via `ggplot2`, it is quite easy to customize parts of the plot beyond some of the arguments that are provided in the `ggcurve()` function. For example, we are able to provide arguments to the function to give custom titles, subtitles, x-axes, y-axes, fills, and colors. However, we could also do this using the standard `ggplot2` grammar. We'll generate a quick graph to show how. 

```{r echo=TRUE, fig.height=4.5, fig.width=6}
library(concurve)
set.seed(1031)
GroupA <- rnorm(500)
GroupB <- rnorm(500)
RandomData <- data.frame(GroupA, GroupB)
intervalsdf <- curve_mean(GroupA, GroupB,
  data = RandomData, method = "default"
)
(function1 <- ggcurve(data = intervalsdf[[1]], type = "c", nullvalue = TRUE))
```

Those are some of the default options provided to `ggcurve()`. We could provide `ggcurve()` arguments for the title, subtitle, etc, but we could also do it like so:

```{r}
library(ggplot2)
function1 +
  labs(
    title = "Random Title",
    subtitle = "Random Subtitle",
    x = "x-axis",
    y = "y-axis",
    caption = "Custom Caption"
  )
```

If we even wanted to provide a custom theme, we could do the following.  

```{r}
library(cowplot)
logo_file <- "https://res.cloudinary.com/less-likely/image/upload/v1575441662/Site/Logo2.jpg"

function1 <- function1 +
  theme_cowplot()

function2 <- ggdraw(function1) +
  draw_image(logo_file, x = 1, y = 1, hjust = 2, vjust = 1.75, width = 0.13, height = 0.2)

function2
```

I've only tried testing this with the `cowplot` package, so I cannot say for sure that the functions won't break when applied with other themes.[@Wilke_2019]

# More Advanced Customization

Although the above shows how flexible concurve's `ggcurve()` is due to the integration with the `ggplot2` framework, we can achieve much more with a newer package called `ggtext`.[@wilkeGgtextImprovedText2020] If we wanted to fully control how the labels and titles in our graphs looked or insert textboxes with full control, we could easily do that with `ggtext.`  

Before we used cowplot, here's what our plain function looked like (after regenerating it from scratch)
```{r}

(function1 <- ggcurve(data = intervalsdf[[1]], type = "c", nullvalue = TRUE))
```
Simple enough, eh?

Okay, but if we wanted to have more fine control over how the title and axes look, we could do that. Here's how:

We take our existing object function and we specify all the usual options for the titles, axes, etc

```{r}
(function1 <- ggcurve(data = intervalsdf[[1]], type = "c", nullvalue = TRUE, title = "Something Super Important", xaxis = "Theta"))
```

But now, instead of doing that above, we'll use a different style to control the size, form, and color of the text. We'll load `ggtext` and then specify the arguments. 

```{r}

library(ggtext)

function1 <- ggcurve(data = intervalsdf[[1]], type = "c", nullvalue = TRUE, title = "Something Super Important", xaxis = "<span style = 'color:#3f8f9b;'>Theta</span> ")

function1 + 
   labs(title = "*P*-value Function / Consonance Curve<br><span style = 'font-size:9pt;'>
The function below contains <span style = 'color:#3f8f9b;'>nested</span> confidence/compatibility intervals at <span style = 'color:#3f8f9b;'>every possible level</span> (95%, 90%, 75%, 50%)
allowing one to see a range of estimates that are <span style = 'color:#3f8f9b;'>consistent with the model and its assumptions</span> that were used to compute
the test statistics, *P*-values, and interval estimates. </span>",
       subtitle = NULL) +
  theme(
    plot.title = element_textbox_simple(
      size = 11, lineheight = 1.1,
      linetype = 1, # turn on border
      box.color = "#748696", # border color
      fill = "white", # background fill color
      r = grid::unit(3, "pt"),
      padding = margin(8, 8, 8, 8), # padding around text inside the box
      maxwidth = unit(8, "in"),# margin outside the box
    ),
    axis.text=element_text(size=10),
    axis.title.y=element_text(size=11),
    axis.title.x = element_textbox_simple(
      size = 10,
      width = NULL,
      padding = margin(4, 4, 4, 4),
      margin = margin(4, 0, 0, 0),
      linetype = 1,
       r = grid::unit(3, "pt"),
      box.color = "#748696", # border color
       fill = "white", # background fill color
    )
  ) 

```

As you might have noticed above, we can control the size of the text, the color of it, whether it's bolded, italicized, etc., and that's partially because it uses markdown. 

# Saving Plots

The most common way useRs save plots is by going to the plots tab in an IDE like RStudio and clicking "export" and then "save as image" or by using `ggsave()`, however, `cowplot` has a superior function with far better default options built into it known as `save_plot()`. 

```{r}
save_plot("function2.pdf", function2)
```

* * * 

# References

* * * 
