---
title: "Constructing Valid Likelihood Functions"
description: "Building exact likelihood functions from first principles for common designs -- proportions, odds ratios, rate ratios, means, and variances -- and coercing them into concurve objects."
output:
  rmarkdown::html_vignette:
    toc: true
opengraph:
  image:
    src: "https://res.cloudinary.com/less-likely/image/upload/v1591475692/Site/lfunctions.png"
  twitter:
    card: summary
    creator: "@dailyzad"
bibliography: references.bib
link-citations: yes
csl: american-medical-association.csl
vignette: >
  %\VignetteIndexEntry{Constructing Valid Likelihood Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  message = TRUE,
  warning = TRUE,
  collapse = TRUE,
  comment = "#>",
  fig.height = 4.5,
  fig.width = 6
)
library(concurve)
```

A likelihood function is only as trustworthy as the model it comes from. It is easy to plot a Gaussian bump centered on an estimate and call it a likelihood; it is harder --- and more honest --- to write down the *exact* likelihood a design implies and plot that.[@royallStatisticalEvidenceLikelihood1997] This vignette builds several likelihood functions from first principles, each valid for its design, and coerces each into the data-frame layout `concurve` uses so that `ggcurve()` can draw it. Where an exact likelihood eliminates a nuisance parameter by conditioning, we use the conditional likelihood rather than a normal approximation.[@coxTheoreticalStatistics1974]

Every construction below is checked: its maximum lands on the estimator the design defines. A standalone script, `verify-likelihoods.R`, repeats those checks as assertions.

# A small helper

Each construction produces a vector of parameter `values` and a relative log-likelihood. This helper normalizes and packages them into the five columns `concurve` expects (`values`, `likelihood`, `loglikelihood`, `support`, `deviancestat`) --- the same object `curve_lik()` returns.

```{r echo=TRUE}
as_concurve_lik <- function(values, loglik) {
  loglik_rel <- loglik - max(loglik) # relative log-likelihood, max = 0
  support <- exp(loglik_rel) # relative likelihood in (0, 1]
  df <- data.frame(
    values        = values,
    likelihood    = support, # relative likelihood      -> type = "l3"
    loglikelihood = loglik_rel, # relative log-likelihood   -> type = "l2"
    support       = support, # relative likelihood       -> type = "l1"
    deviancestat  = -loglik_rel # -log relative likelihood  -> type = "d"
  )
  class(df) <- c("data.frame", "concurve")
  df
}

# collect every rho whose relative likelihood clears the 1/k cutoff
support_interval <- function(df, k) {
  keep <- df$values[df$support >= 1 / k]
  c(lower = min(keep), upper = max(keep))
}
```

# 1. A single proportion: the exact binomial likelihood

With $x$ successes in $n$ independent trials, the likelihood of the risk $p$ is binomial, $L(p) \propto p^{x}(1-p)^{n-x}$, and its maximum is the sample proportion $\hat p = x/n$. No approximation is needed --- this is the likelihood.

```{r echo=TRUE}
x <- 8
n <- 20
phat <- x / n

p <- seq(1e-4, 1 - 1e-4, length.out = 4000)
loglik_binom <- x * log(p) + (n - x) * log(1 - p)
lik_p <- as_concurve_lik(p, loglik_binom)

ggcurve(lik_p,
  type = "l1", nullvalue = TRUE,
  xaxis = "Risk (p)",
  title = "Exact Binomial Likelihood",
  subtitle = sprintf("x = %d successes in n = %d trials", x, n)
)
```

## Why "exact" earns its keep

The Wald approximation replaces this with a symmetric Gaussian of variance $\hat p(1-\hat p)/n$. Near the boundaries the two disagree visibly --- the exact likelihood is skewed and respects $0 < p < 1$, while the Wald bump does not.[@rothmanModernEpidemiology2008]

```{r echo=TRUE}
se_wald <- sqrt(phat * (1 - phat) / n)
loglik_wald <- -0.5 * (p - phat)^2 / se_wald^2
lik_p_wald <- as_concurve_lik(p, loglik_wald)

ggcurve(lik_p_wald,
  type = "l1", nullvalue = TRUE,
  xaxis = "Risk (p)",
  title = "Wald (Normal) Approximation",
  subtitle = "Symmetric by construction -- compare the skew above"
)
```

```{r echo=TRUE}
rbind(
  exact = support_interval(lik_p, 6.8),
  wald  = support_interval(lik_p_wald, 6.8)
)
```

The exact interval is the one to report; the Wald interval is what happens when you let algebra outrun the data.

# 2. An odds ratio: the exact conditional likelihood

For a $2 \times 2$ table, conditioning on both margins removes the nuisance baseline odds and leaves an **exact** likelihood for the odds ratio $\psi$ --- the noncentral hypergeometric.[@coxTheoreticalStatistics1974] If the exposed-case count is $a$, with exposed total $n_1$, unexposed total $n_0$, and case total $m_1$,

$$\ell(\psi) = a\log\psi \;-\; \log\!\sum_{k}\binom{n_1}{k}\binom{n_0}{m_1-k}\psi^{k},$$

summed over the feasible range of $k$. We plot on the log-OR scale, where the null is $0$.

```{r echo=TRUE}
# table:            D+   D-
#         E+         a    b
#         E-         c    d
a <- 12
b <- 8
c <- 5
d <- 15
n1 <- a + b
n0 <- c + d
m1 <- a + c

cond_loglik_or <- function(logpsi) {
  klo <- max(0, m1 - n0)
  khi <- min(m1, n1)
  k <- klo:khi
  logterms <- lchoose(n1, k) + lchoose(n0, m1 - k) + k * logpsi
  M <- max(logterms)
  logZ <- M + log(sum(exp(logterms - M))) # log-sum-exp for stability
  a * logpsi - logZ
}

logpsi <- seq(-3, 4, length.out = 4000)
loglik_or <- vapply(logpsi, cond_loglik_or, numeric(1))
lik_or <- as_concurve_lik(logpsi, loglik_or)

ggcurve(lik_or,
  type = "l1", nullvalue = 0,
  xaxis = "log(Odds Ratio)",
  title = "Exact Conditional Likelihood for the Odds Ratio",
  subtitle = "Noncentral hypergeometric -- free of the baseline odds"
)

ggcurve(lik_or,
  type = "d", nullvalue = 0,
  xaxis = "log(Odds Ratio)",
  title = "Deviance Function"
)
```

Note the peak: the conditional likelihood is maximized at the *conditional* MLE, which is close to but **not identical** to the naive cross-product $ad/bc$.

```{r echo=TRUE}
c(
  `sample ad/bc` = (a * d) / (b * c),
  `conditional MLE` = exp(lik_or$values[which.max(lik_or$support)])
)
```

That gap is not rounding error --- it is the difference between a marginal summary and the estimator the exact conditional model actually supports.[@coxTheoreticalStatistics1974]

# 3. A rate ratio: the exact conditional binomial likelihood

Two Poisson counts, $a$ events in person-time $T_1$ and $b$ events in $T_0$, carry an exact likelihood for the rate ratio $\theta = \lambda_1/\lambda_0$ once you condition on the total number of events. Conditionally, $a \mid (a+b) \sim \text{Binomial}(a+b,\ \pi)$ with $\pi(\theta) = \theta T_1 / (\theta T_1 + T_0)$.[@cummingsAnalysisIncidenceRates2019]

```{r echo=TRUE}
a_e <- 30
T1 <- 1000 # exposed:   events, person-time
b_e <- 18
T0 <- 1200 # unexposed: events, person-time

cond_loglik_rr <- function(logtheta) {
  theta <- exp(logtheta)
  pi <- theta * T1 / (theta * T1 + T0)
  a_e * log(pi) + b_e * log(1 - pi)
}

logtheta <- seq(-2, 3, length.out = 4000)
loglik_rr <- vapply(logtheta, cond_loglik_rr, numeric(1))
lik_rr <- as_concurve_lik(logtheta, loglik_rr)

ggcurve(lik_rr,
  type = "l1", nullvalue = 0,
  xaxis = "log(Rate Ratio)",
  title = "Exact Conditional Likelihood for the Rate Ratio",
  subtitle = "Binomial conditional on total events -- no offset approximation"
)
```

Here the conditional MLE *is* the sample rate ratio exactly:

```{r echo=TRUE}
c(
  `sample RR` = (a_e / T1) / (b_e / T0),
  `likelihood peak` = exp(lik_rr$values[which.max(lik_rr$support)])
)
```

# 4. A mean: the profile likelihood

For a normal sample, profiling out the unknown variance leaves a profile likelihood for the mean $\mu$ that depends only on the residual sum of squares (Pawitan, *In All Likelihood*, 2001):[@royallStatisticalEvidenceLikelihood1997]

$$\ell_p(\mu) = -\frac{n}{2}\log\!\left(\frac{\sum_i (x_i - \mu)^2}{\sum_i (x_i - \bar x)^2}\right).$$

```{r echo=TRUE}
xs <- c(5.1, 4.8, 5.5, 4.9, 5.3, 5.0, 4.7, 5.2, 5.4, 4.6)
n_x <- length(xs)
xbar <- mean(xs)
ss_min <- sum((xs - xbar)^2)

mu <- seq(4.4, 5.8, length.out = 4000)
loglik_mu <- -(n_x / 2) * log(vapply(mu, function(m) sum((xs - m)^2), numeric(1)) / ss_min)
lik_mu <- as_concurve_lik(mu, loglik_mu)

ggcurve(lik_mu,
  type = "l1", nullvalue = 5.0,
  xaxis = expression(mu),
  title = "Profile Likelihood for the Mean"
)
```

# 5. A variance (and SD): the chi-square likelihood

The likelihood for a variance $\sigma^2$ follows from the sampling distribution of $s^2$, since $\nu s^2 / \sigma^2 \sim \chi^2_\nu$ with $\nu = n-1$. It is genuinely asymmetric --- a shape people rarely see plotted because the Wald habit hides it. We plot it on the standard- deviation scale by mapping $\sigma = \sqrt{\sigma^2}$.

```{r echo=TRUE}
s2 <- var(xs)
nu <- n_x - 1

sig2 <- seq(0.02, 0.60, length.out = 4000)
loglik_var <- -(nu / 2) * log(sig2) - nu * s2 / (2 * sig2)
lik_sd <- as_concurve_lik(sqrt(sig2), loglik_var) # values on the SD scale

ggcurve(lik_sd,
  type = "l1", nullvalue = TRUE,
  xaxis = expression(sigma),
  title = "Likelihood for the Standard Deviation",
  subtitle = "Asymmetric by nature -- the chi-square shows through"
)
```

# What made each of these valid

None of these leaned on a Gaussian approximation to the log-likelihood. The proportion used the binomial directly; the odds ratio and rate ratio used exact conditional likelihoods that delete their nuisance parameters by conditioning on sufficient margins; the mean and variance came from the exact normal-theory sampling distributions.[@coxTheoreticalStatistics1974; @royallStatisticalEvidenceLikelihood1997] Each maximum coincides with the estimator its design defines (the odds-ratio *conditional* MLE included), and each function is skewed exactly as much as the model says it should be --- which is the information a single interval throws away.[@rafiSemanticCognitiveTools2020; @pooleConfidenceIntervalsExclude1987]

Once an object is in `concurve` form, everything else in the package --- support intervals, deviance and log-likelihood views, side-by-side comparison with a consonance function --- works on it unchanged.

# Cite R Packages

Please remember to cite the packages that you use.

```{r}
citation("concurve")
```

# References

------------------------------------------------------------------------
