---
title: "Constructing Likelihood Functions with Existing R Tools"
description: "Any tool that returns a (profile) log-likelihood -- optim, maxLik, bbmle, ProfileLikelihood, cond -- can feed a concurve likelihood object through one small adapter."
output:
  rmarkdown::html_vignette:
    toc: true
opengraph:
  image:
    src: "https://res.cloudinary.com/less-likely/image/upload/v1591475692/Site/lfunctions.png"
  twitter:
    card: summary
    creator: "@dailyzad"
bibliography: references.bib
link-citations: yes
csl: american-medical-association.csl
vignette: >
  %\VignetteIndexEntry{Constructing Likelihood Functions with Existing R Tools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  message = TRUE,
  warning = TRUE,
  collapse = TRUE,
  comment = "#>",
  fig.height = 4.5,
  fig.width = 6
)
library(concurve)

# availability guards -- these packages are illustrative and are NOT concurve
# dependencies, so every chunk that uses one is gated on it being installed.
has_maxLik <- requireNamespace("maxLik", quietly = TRUE)
has_bbmle <- requireNamespace("bbmle", quietly = TRUE)
has_cond <- requireNamespace("cond", quietly = TRUE)
```

R has no shortage of machinery for building likelihoods: `optim()` and
`optimize()` in base R, `maxLik`, `bbmle::mle2()`, `ProfileLikelihood`,
and `cond` for exact conditional GLM
inference.[@choiProfileLikelihoodProfileLikelihood2011] They look
different, but for `concurve` they all reduce to the same thing: a grid
of parameter `values` and a **profile** log-likelihood evaluated on it.
This vignette shows several of those tools feeding one adapter.

> Note: `maxLik`, `bbmle`, and `cond` are not `concurve` dependencies,
> so the chunks that use them run only if you have them installed. The
> `optim()` route is base R and always runs.

# One adapter for all of them

Whatever produces the log-likelihood, this helper normalizes it and
packs it into the five columns `concurve` uses (`values`, `likelihood`,
`loglikelihood`, `support`, `deviancestat`) --- the object `curve_lik()`
returns and `ggcurve()` draws.

```{r echo=TRUE}
as_concurve_lik <- function(values, loglik) {
  loglik_rel <- loglik - max(loglik)
  support <- exp(loglik_rel)
  df <- data.frame(
    values        = values,
    likelihood    = support,
    loglikelihood = loglik_rel,
    support       = support,
    deviancestat  = -loglik_rel
  )
  class(df) <- c("data.frame", "concurve")
  df
}
```

The only rule that matters for validity: when the model has more than
one parameter, the log-likelihood you evaluate at each grid point must
be **profiled** over the others (maximized over the nuisance
parameters), not merely sliced at their
estimates.[@coxTheoreticalStatistics1974] We return to that at the end,
because it is the single most common way a "likelihood" plot goes wrong.

Some data to reuse:

```{r echo=TRUE}
set.seed(123)
x <- rnorm(50, mean = 9.3, sd = 3.2)
c(mean = mean(x), sd = sd(x))
```

# Route 1: base R `optim()` (+ observed-information SEs)

Minimize the negative log-likelihood to get the MLE and, from the
Hessian, the observed-information standard errors --- exactly the
pattern in the pasted script.

```{r echo=TRUE}
nll_normal <- function(theta, x) {
  mu <- theta[1]
  sigma <- theta[2]
  -sum(dnorm(x, mean = mu, sd = sigma, log = TRUE))
}

fit <- optim(c(mu = 8, sigma = 3), nll_normal,
  x = x,
  method = "BFGS", hessian = TRUE
)
fit$par
sqrt(diag(solve(fit$hessian))) # SE(mu), SE(sigma)
```

To turn this into a likelihood *function* for the mean, profile out
`sigma` at each candidate `mu` --- and here profiling is free, since the
conditional MLE of `sigma` given `mu` is closed-form.

```{r echo=TRUE}
n <- length(x)
mu_grid <- seq(fit$par["mu"] - 3, fit$par["mu"] + 3, length.out = 4000)
loglik_mu <- vapply(mu_grid, function(m) {
  sigma2_hat <- mean((x - m)^2) # profile: MLE of sigma^2 given mu
  -(n / 2) * (log(2 * pi) + log(sigma2_hat) + 1)
}, numeric(1))

lik_optim <- as_concurve_lik(mu_grid, loglik_mu)

ggcurve(lik_optim,
  type = "l1", nullvalue = round(fit$par["mu"], 2),
  xaxis = expression(mu),
  title = "Likelihood for the Mean (via optim profiling)"
)
```

# Route 2: `maxLik`

`maxLik` maximizes a log-likelihood you supply (returning the value, not
its negative) and gives a tidy `summary()` with SEs. To get the function
rather than just the point, evaluate the profiled log-likelihood on a
grid and hand it to the adapter.

```{r eval=has_maxLik, echo=TRUE}
library(maxLik)

loglik_normal <- function(theta, x) {
  sum(dnorm(x, mean = theta[1], sd = sqrt(theta[2]), log = TRUE))
}
start <- c(mu = 8, sig2 = 9)
theta_mle <- maxLik(loglik_normal, start = start, x = x)
summary(theta_mle)

# profile likelihood for mu, reusing the closed-form sigma^2 profile
loglik_mu2 <- vapply(mu_grid, function(m) {
  loglik_normal(c(m, mean((x - m)^2)), x = x)
}, numeric(1))

ggcurve(as_concurve_lik(mu_grid, loglik_mu2),
  type = "l1",
  nullvalue = round(coef(theta_mle)[1], 2),
  xaxis = expression(mu),
  title = "Likelihood for the Mean (via maxLik)"
)
```

```{r eval=!has_maxLik, echo=FALSE, results="asis"}
cat("> `maxLik` is not installed, so this route is not evaluated. Install it with `install.packages(\"maxLik\")` to run it.")
```

# Route 3: `bbmle::mle2()` with built-in profiling

`bbmle` is purpose-built for this: `mle2()` fits by maximum likelihood
and `profile()` returns the profile likelihood for each parameter, which
is exactly what we want. `bbmle` reports the profile on the **signed
square-root deviance** ($z$) scale; squaring recovers the deviance, and
$\exp(-\tfrac12 z^2)$ recovers the relative likelihood.

```{r eval=has_bbmle, echo=TRUE}
library(bbmle)

nll <- function(mu, sigma) -sum(dnorm(x, mean = mu, sd = sigma, log = TRUE))
m <- mle2(nll, start = list(mu = 8, sigma = 3), data = list(x = x))

pr <- profile(m)
pr_mu <- as.data.frame(pr) # columns: param, z, par.vals...
pr_mu <- pr_mu[pr_mu$param == "mu", ]

mu_vals <- pr_mu[["par.vals.mu"]]
if (is.null(mu_vals)) mu_vals <- pr_mu[["mu"]] # column name varies by version

lik_bbmle <- as_concurve_lik(mu_vals, -0.5 * pr_mu$z^2) # loglik_rel = -z^2/2

ggcurve(lik_bbmle,
  type = "l1", nullvalue = round(coef(m)["mu"], 2),
  xaxis = expression(mu),
  title = "Profile Likelihood for the Mean (via bbmle)"
)
```

```{r eval=!has_bbmle, echo=FALSE, results="asis"}
cat("> `bbmle` is not installed, so this route is not evaluated. Install it with `install.packages(\"bbmle\")` to run it.")
```

# Route 4: exact conditional GLM inference with `cond`

For a logistic model, `cond` computes higher-order **exact conditional**
likelihood inference for a coefficient of interest, eliminating the
nuisance parameters by conditioning rather than by asymptotic
approximation.[@coxTheoreticalStatistics1974] This is the same spirit as
the exact conditional odds-ratio likelihood in the *Constructing Valid
Likelihood Functions* vignette, but for a full regression.

```{r eval=has_cond, echo=TRUE}
library(cond)
data(babies)

mod <- glm(cbind(r1, r2) ~ day + lull - 1, family = binomial, data = babies)
cond_fit <- cond(mod, offset = lullyes) # exact conditional inference for 'lull'
summary(cond_fit)
plot(cond_fit, which = 2) # conditional (modified profile) likelihood
```

```{r eval=!has_cond, echo=FALSE, results="asis"}
cat("> `cond` is not installed, so this route is not evaluated. Install it with `install.packages(\"cond\")` to run it. Its `plot()` draws the conditional and modified-profile likelihoods directly; to bring them into concurve, pass the object's grid of parameter values and conditional log-likelihood to `as_concurve_lik()`.")
```

`cond` plots the conditional and modified-profile likelihoods itself; to
render them in `concurve`, extract its grid of values and conditional
log-likelihood and pass them to `as_concurve_lik()` as above.

# The pitfall: slicing is not profiling

The pasted script contains a `likelihood.normal.mu(mu, sig2 = 1, x)`
that fixes $\sigma^2 = 1$. That is a **slice** through the likelihood at
an assumed variance, not a profile. It peaks in the right place but
carries the wrong curvature --- so every interval read off it is wrong
unless $\sigma^2$ really is 1.

```{r echo=TRUE}
loglik_slice <- vapply(mu_grid, function(m) {
  -0.5 * sum((x - m)^2) # fixes sigma^2 = 1
}, numeric(1))

lik_slice <- as_concurve_lik(mu_grid, loglik_slice)

ggcurve(lik_slice,
  type = "l1", nullvalue = round(fit$par["mu"], 2),
  xaxis = expression(mu),
  title = "WRONG: mean likelihood with sigma^2 fixed at 1",
  subtitle = "Too narrow -- the real sigma is about 3.2, not 1"
)
```

Compare its width to the profiled version from Route 1: the slice is far
too tight because it pretends the variance is known and small. When you
build a likelihood with any of the tools above, profile the nuisance
parameters --- `optim()` over them, `bbmle::profile()`, `maxLik` on a
profiled objective, or a package like `cond` that conditions them away
exactly.[@royallStatisticalEvidenceLikelihood1997;
@coxTheoreticalStatistics1974]

# Likelihood meets the P-value function

The payoff of building likelihoods properly is that they line up with
the consonance (P-value) function for the same parameter --- the two are
coordinate views of one inferential object. Here we make that visible
with `plot_compare()` using the native constructors `curve_lik_glm()`
and `curve_analytic()` (this section evaluates when those functions are
available in your installed version of `concurve`).

```{r echo=FALSE}
has_native <- exists("curve_lik_glm") && exists("curve_analytic") &&
  exists("curve_support") && exists("as_curve_lik")
```

```{r eval=has_native, echo=TRUE}
mod <- glm(am ~ mpg, family = binomial, data = mtcars)
est <- coef(mod)["mpg"]
se <- sqrt(diag(vcov(mod)))["mpg"]

# profile likelihood: nuisance intercept maximized out at every grid point
lik_prof <- curve_lik_glm(mod, "mpg", steps = 200)

# Wald likelihood: the normal approximation implied by (est, se) --
# exactly the likelihood the consonance function from curve_analytic encodes
grid <- lik_prof[[1]]$values
lik_wald <- as_curve_lik(grid, dnorm(grid, mean = est, sd = se, log = TRUE))

plot_compare(lik_prof[[1]], lik_wald[[1]],
  type = "l1",
  title = "Profile vs Wald Likelihood",
  subtitle = "Same coefficient; the gap between them is the skew the Wald approximation ignores",
  xaxis = expression(beta[mpg])
)
```

The two curves agree near the maximum and part company in the shoulders
--- which is precisely where interval limits live. The numeric version
of the same statement: the 1/6.83 support interval from the profile
likelihood reproduces the 95% consonance interval from the profile-based
function, while the Wald consonance interval matches the Wald
likelihood.

```{r eval=has_native, echo=TRUE}
# 95% consonance interval, Wald construction
cons <- curve_analytic(estimate = est, se = se, dist = "z")
ci95 <- cons[[1]][
  which.min(abs(cons[[1]]$intrvl.level - 0.95)),
  c("lower.limit", "upper.limit")
]

rbind(
  `profile 1/6.83 support` = unlist(curve_support(lik_prof[[1]], 6.83)[c("lower.limit", "upper.limit")]),
  `Wald    1/6.83 support` = unlist(curve_support(lik_wald[[1]], 6.83)[c("lower.limit", "upper.limit")]),
  `Wald    95% consonance` = unlist(ci95)
)
```

The Wald support interval and the Wald consonance interval are identical
by construction ($e^{-z^2/2}$ at $z = 1.96$ is $1/6.83$); the profile
versions differ from both exactly as much as the likelihood is skewed.
Report the profile-based ones.

```{r eval=!has_native, echo=FALSE, results="asis"}
cat("> This section requires a version of `concurve` that includes the native likelihood constructors (`curve_lik_glm()`, `curve_analytic()`, `curve_support()`, `as_curve_lik()`).")
```

# Adding these tools to your package

If you want any of the guarded routes to run when the vignettes build on
CRAN, add the packages to `Suggests` in `DESCRIPTION` (`maxLik`,
`bbmle`, `cond`) and keep the `requireNamespace()` guards --- that is
the pattern CRAN expects for optional, illustrative dependencies.

# Cite R Packages

Please remember to cite the packages that you use.

```{r}
citation("concurve")
```

# References

------------------------------------------------------------------------
