---
title: "Logistic Regression in R"
description: "A simple introduction to using the concurve R package for logistic regression."
output: 
  rmarkdown::html_vignette:
    toc: true
opengraph:
  image: 
    src: "https://res.cloudinary.com/less-likely/image/upload/v1554700143/Site/Projects.jpg"
  twitter:
    card: summary
    creator: "@dailyzad"    
bibliography: references.bib
link-citations: yes
csl: american-medical-association.csl
vignette: >
  %\VignetteIndexEntry{Logistic Regression in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Suppose we wanted to produce confidence distributions for data with binary outcomes and where we employ a logistic regression, we would do the following. Here, I use the mtcars dataset for the example and also simulate some very simple binary data. We use `suppressMessages()` to avoid seeing the long list of profiling messages.

```{r}
library(concurve)

X <- rnorm(100, mean = 0, sd = 1)
Y <- rbinom(n = 100, size = 1, prob = 0.5)

mydata1 <- data.frame(X, Y)


model1 <- glm(Y ~ X,
  data = mydata1,
  family = binomial(link = "logit"))

model2 <- glm(am ~ mpg, family = binomial(link = "logit"), data = mtcars)

summary(model1)
summary(model2)



model_pro <- suppressMessages(curve_gen(
  model = model1,
  var = "X",
  method = "glm",
  log = T,
  steps = 1000,
  table = TRUE))

model_con <- suppressMessages(curve_gen(
  model = model2,
  var = "mpg",
  method = "glm",
  log = T,
  steps = 1000,
  table = TRUE))

head(model_con[[1]], 10)


(ggcurve(model_con[[1]],
    measure = "ratio",
    type = "c",
    nullvalue = c(0.8, 1.2), title = "Confidence Curve",
    subtitle = "The function displays intervals at every level.",
    xaxis = expression(theta == ~"Range of Values"),
    yaxis1 = expression(paste(italic(p), "-value")),
    yaxis2 = "Levels for CI (%)"))


(ggcurve(model_pro[[1]],
    measure = "ratio",
    type = "c",
    nullvalue = c(0.8, 1.2), title = "Confidence Curve",
    subtitle = "The function displays intervals at every level.",
    xaxis = expression(theta == ~"Range of Values"),
    yaxis1 = expression(paste(italic(p), "-value")),
    yaxis2 = "Levels for CI (%)"))

(ggcurve(model_con[[2]],
    measure = "ratio",
    type = "cdf",
    nullvalue = c(0.8, 1.2), title = "Confidence Distribution",
    subtitle = "The function displays intervals at every level.",
    xaxis = expression(theta == ~"Range of Values"),
    yaxis1 = expression(paste(italic(p), "-value")),
    yaxis2 = "Levels for CI (%)"))

(ggcurve(model_pro[[2]],
    measure = "ratio",
    type = "cdf",
    nullvalue = c(0.8, 1.2), title = "Confidence Distribution",
    subtitle = "The function displays intervals at every level.",
    xaxis = expression(theta == ~"Range of Values"),
    yaxis1 = expression(paste(italic(p), "-value")),
    yaxis2 = "Levels for CI (%)"))


(ggcurve(model_con[[2]],
    measure = "ratio",
    type = "cd",
    nullvalue = NULL, title = "Confidence Density",
    subtitle = "The function displays intervals at every level.",
    xaxis = expression(theta == ~"Range of Values"),
    yaxis1 = expression(paste(italic(p), "-value")),
    yaxis2 = "Levels for CI (%)"))

(ggcurve(model_pro[[2]],
    measure = "ratio",
    type = "cd",
    nullvalue = NULL, title = "Confidence Density",
    subtitle = "The function displays intervals at every level.",
    xaxis = expression(theta == ~"Range of Values"),
    yaxis1 = expression(paste(italic(p), "-value")),
    yaxis2 = "Levels for CI (%)"))

```

# Cite R Packages

Please remember to cite the R packages that you use in your work.

```{r}
citation("concurve")
citation("cowplot")
```

* * * 

# References

* * * 