---
title: "Count and Rate Models: Likelihood and P-value Functions"
description: "Constructing a profile likelihood function for a Poisson rate ratio with curve_lik(), then pairing it with the consonance (P-value) function from curve_gen()."
output:
  rmarkdown::html_vignette:
    toc: true
opengraph:
  image:
    src: "https://res.cloudinary.com/less-likely/image/upload/v1591475692/Site/lfunctions.png"
  twitter:
    card: summary
    creator: "@dailyzad"
bibliography: references.bib
link-citations: yes
csl: american-medical-association.csl
vignette: >
  %\VignetteIndexEntry{Count and Rate Models: Likelihood and P-value Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  message = TRUE,
  warning = TRUE,
  collapse = TRUE,
  comment = "#>",
  fig.height = 4.5,
  fig.width = 6
)
library(concurve)
```

Count outcomes --- events per person, per unit time, per unit area --- are the natural home of the Poisson model, and the parameter that epidemiologists actually care about is the **rate ratio**.[@cummingsAnalysisIncidenceRates2019] This vignette leads with the *likelihood function* for a Poisson log-rate-ratio and then shows the matching consonance (P-value) function, so you can see the same coefficient described two ways.[@rafiSemanticCognitiveTools2020; @fraserPvalueFunctionStatistical2019]

# The data

We simulate counts for an exposed vs. unexposed group, adjusting for one continuous covariate. The true log-rate-ratio for `group` is $0.8$ (a rate ratio of about $2.2$).

```{r echo=TRUE}
set.seed(2718)
n <- 200
x1 <- rnorm(n)
group <- rbinom(n, size = 1, prob = 0.5)
eta <- -0.5 + 0.8 * group + 0.3 * x1
y <- rpois(n, lambda = exp(eta))

pois_dat <- data.frame(y, group, x1)
head(pois_dat)
```

# Constructing the likelihood function

We reuse the same machinery as the profile-likelihood vignette: `ProfileLikelihood::profilelike.glm()` profiles the Poisson GLM over the nuisance coefficients at each value of the target parameter, and `curve_lik()` turns that into a `concurve` likelihood object.[@choiProfileLikelihoodProfileLikelihood2011; @coleMaximumLikelihoodProfile2013] The variable named in `profile.theta` (here `group`) is the one profiled; it is left out of the formula.

```{r echo=TRUE}
library(ProfileLikelihood)

xx <- profilelike.glm(
  y ~ x1,
  data          = pois_dat,
  profile.theta = "group",
  family        = poisson(link = "log"),
  length        = 500,
  round         = 2
)

lik <- curve_lik(xx, data = pois_dat)
```

The profiled parameter is on the **log scale**, so `values` here are candidate log-rate-ratios and the null of "no association" is at $0$.

## Plotting the likelihood

```{r echo=TRUE}
ggcurve(lik[[1]],
  type = "l1", nullvalue = TRUE,
  xaxis = "log(Rate Ratio)",
  title = "Relative Likelihood Function",
  subtitle = "Profile likelihood for the Poisson log-rate-ratio"
)

ggcurve(lik[[1]],
  type = "l2",
  xaxis = "log(Rate Ratio)",
  title = "Log-Likelihood Function"
)

ggcurve(lik[[1]],
  type = "d", nullvalue = 0,
  xaxis = "log(Rate Ratio)",
  title = "Deviance Function"
)
```

The likelihood peaks near the true value of $0.8$, and the deviance function --- $-\log$ relative likelihood --- turns the same curve into a "distance from the best-supported value" scale, where Royall's $1/8$ and $1/32$ support cutoffs correspond to deviance heights of about $2.08$ and $3.47$.[@royallStatisticalEvidenceLikelihood1997]

# The companion: consonance and surprisal functions

For the frequentist view we fit the full Poisson GLM and hand it to `curve_gen()` with `method = "glm"`. Setting `log = TRUE` keeps the underlying values on the log scale but lets us plot on the **rate-ratio** scale with `measure = "ratio"`, where the null is $1$.[@rafiSemanticCognitiveTools2020] We wrap the call in `suppressMessages()` to hide the long list of profiling messages.

```{r echo=TRUE}
pois_mod <- glm(y ~ group + x1, data = pois_dat, family = poisson(link = "log"))

rate_curve <- suppressMessages(curve_gen(
  model  = pois_mod,
  var    = "group",
  method = "glm",
  log    = TRUE,
  steps  = 1000
))
```

```{r echo=TRUE}
ggcurve(rate_curve[[1]],
  type = "c", measure = "ratio", nullvalue = 1,
  xaxis = "Rate Ratio",
  title = "Consonance (P-value) Function"
)

ggcurve(rate_curve[[1]],
  type = "s", measure = "ratio", nullvalue = 1,
  xaxis = "Rate Ratio",
  title = "Surprisal (S-value) Function"
)
```

The consonance function peaks at the estimated rate ratio; its slice at $1 - \alpha = 0.95$ is the ordinary 95% rate-ratio interval. The surprisal function reports $-\log_2 p$ against each candidate rate ratio, so a rate ratio of $1$ carries a readable number of **bits** of information against it.[@rafiSemanticCognitiveTools2020; @greenlandStatisticalTestsValues2016]

## A confidence distribution too

Because `curve_gen()` also returns the density-function values, you can draw the full confidence distribution and density for the rate ratio.[@schwederConfidenceLikelihood2002; @xieConfidenceDistributionFrequentist2013]

```{r echo=TRUE}
ggcurve(rate_curve[[2]],
  type = "cdf", measure = "ratio", nullvalue = 1,
  xaxis = "Rate Ratio",
  title = "Consonance Distribution"
)

ggcurve(rate_curve[[2]],
  type = "cd", measure = "ratio",
  xaxis = "Rate Ratio",
  title = "Consonance Density"
)
```

# Putting it together

The log-rate-ratio likelihood and the rate-ratio consonance function are the same coefficient viewed on two scales and through two inferential lenses. The likelihood asks *how well each value is supported by the data*; the consonance function asks *how compatible each value is under the model*.[@rafiSemanticCognitiveTools2020] For a smooth one-parameter target like this, the $1/6.8$ likelihood interval and the 95% consonance interval will land in almost the same place --- which is exactly why neither a single interval nor a single P-value deserves a bright line.[@pooleConfidenceIntervalsExclude1987]

# Cite R Packages

Please remember to cite the packages that you use.

```{r}
citation("concurve")
citation("ProfileLikelihood")
```

# References

------------------------------------------------------------------------
