---
title: "Generating Tables"
description: "This vignette walks through how to construct tables from the curve functions."
output: rmarkdown::html_vignette
opengraph:
  image: 
    src: "https://res.cloudinary.com/less-likely/image/upload/v1591475692/Site/lfunctions.png"
  twitter:
    card: summary
    creator: "@dailyzad"
bibliography: references.bib
link-citations: yes
csl: american-medical-association.csl
vignette: >
  %\VignetteIndexEntry{Generating Tables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  message = TRUE,
  warning = TRUE,
  collapse = TRUE,
  comment = "#>"
)
```

It is quite easy to generate tables of outputs for the various `concurve` functions and in different formats. Here we show how to do this with a simple example. First, we'll simulate some fake data as usual, compare the means, and then produce a confidence function of the outputs.  

```{r echo=TRUE, fig.height=4.5, fig.width=6}
library(concurve)
GroupA <- rnorm(500)
GroupB <- rnorm(500)
RandomData <- data.frame(GroupA, GroupB)

intervalsdf <- curve_mean(GroupA, GroupB,
  data = RandomData, method = "default"
)
```

The results are now stored in intervalsdf. Using the curve_table() function, we can now produce a high quality table with several values of interest.

```{r echo=TRUE, fig.height=2, fig.width=4}
(x <- curve_table(data = intervalsdf[[1]], format = "image"))
```

Here we specified the format as "image", which will give us just that. We can also specify other options such as 

```{r echo=TRUE, fig.height=2, fig.width=4}
(z <- curve_table(intervalsdf[[1]], format = "latex"))
```

which is useful for inserting the output into a TeX document, and we can also specify options such as 

```{r echo=TRUE, fig.height=2, fig.width=4}
(df <- curve_table(intervalsdf[[1]], format = "data.frame"))
```

The options "pptx" and "docx" can also be specified as format options, but specifying these will open those programs if they are installed. 

# Session info

```{r session_info, include=TRUE, echo=FALSE}
sessionInfo()
```

* * * 

# References

* * * 