---
title: "Palmer Penguins"
author: "Cobblepot Analytics"
output: rmarkdown::html_vignette
---

```{r setup, include = FALSE}
req_packs <- c(
  "rms", "mice", "magrittr", "tidyr", "future", "future.apply", "showtext", "concurve", "ggtext", "tidyverse",           "rstan", "latex2exp", "MCMCpack", "quantreg",     "miceadds", "missForest", "randomForest",               "miceFast", "parallelly", "checkmate", "posterior", "kableExtra", "reticulate", "broom.mixed", "xfun", "mgcv",
  "ggplot2", "parallel", "kableExtra", "cowplot", "Cairo", "svglite", "yardstick", "broom", "mi", "loo", "progress",
  "Hmisc", "here", "VIM", "wesanderson", "bayesplot", "lattice", "doParallel", "blogdown", "Amelia", "pbmcapply", "mitml",
  "ggcorrplot", "qqplotr", "car", "Statamarkdown", "rstan", "missForest", "htmltools", "tidybayes", "StanHeaders",
  "summarytools", "performance", "mvtnorm", "ImputeRobust", "gamlss", "ProfileLikelihood", "reshape2", "gtsummary",
  "boot", "knitr", "MASS",    "bootImpute", "brms", "rmarkdown", "tinytex", "texPreview", "pbmcapply", "cli"
)

suppressMessages(lapply(req_packs, library,
                        character.only = TRUE
))
options(parallelly.makeNodePSOCK.setup_strategy = "sequential")
options(parallelly.availableCores.methods = "mc.cores")

```

```{r}
require(Hmisc)
require(rms)   # Version 8.1-0 is required
require(ggplot2)
options(parallelly.makeNodePSOCK.setup_strategy = "sequential")
options(parallelly.availableCores.methods = "mc.cores")

```

```{r}
simd <- function(n, p) {
  x <- matrix(rnorm(n * p), n, p)
  L <- x[, 1]
  y <- ifelse(runif(n) <= plogis(L), 1, 0)
  list(x=x, y=y)
}
```

In what follows let $n=200$ be the sample size and $p=15$ be the number
of predictors. This is a setting with major overfitting.

To check confidence interval coverage we need to know the true Brier
score for our data generating model. This can be well estimated by
simulating a single sample of 200,000 subjects, and computing the Brier
score for many models fitted on 200 subjects, averaging their Brier
scores.

```{r}
set.seed(1)
N    <- 200
np   <- 15
Nbig <- 200000
big  <- simd(Nbig, np)
big  <- data.frame(x = I(big$x), y=big$y)   # I keeps matrix x as a matrix
Y    <- big$y
nsim <- 5000
```

Simulate `r nsim` samples of size 200. For each sample run 300 bootstrap
resamples to compute approximate confidence limits. Also compute the
true performance measures for each sample, to average.

```{r}
# Define a function to generate and analyze one dataset
sim1 <- function(n, p, B=300) {
  .i. <<- .i. + 1
  cat(.i., '', file='/tmp/z', append=TRUE)
  d    <- simd(n, np)
  f    <- lrm(y ~ x, data=d, x=TRUE, y=TRUE)
  phat <- predict(f, newdata=big, type='fitted')

  trueB     <- mean((phat - Y) ^ 2)
  trueDxy   <- somers2(phat, Y)['Dxy']
  trueSlope <- lrm.fit(qlogis(phat), Y)$coef[2]
  # saveraw is passed to predab.re-sample in rms 8.1-0 to cause indexes
  # from individual bootstrap resamples to be saved in a global object .predab_raw.
  v <- validate(f, B=B, saveraw=TRUE)
  r <- .predab_raw.
  w <- which(names(r$orig) %in% c('Dxy', 'Slope', 'B'))
  r$orig   <- r$orig[w]
  r$btrain <- r$btrain[, w]
  r$btest  <- r$btest [, w]
  praw[[.i.]] <<- r
  v <- v[w, c('index.corrected', 'Lower', 'Upper')]
  v <- cbind(v, true=c(trueDxy, trueSlope, trueB))
  v
}
.i.  <- 0
praw <- NULL
set.seed(1)
if(file.exists('indexes.rds')) {
  s    <- readRDS('indexes.rds')
  praw <- readRDS('praw.rds')
} else {
  s <- replicate(nsim, sim1(n=N, p=np))
  saveRDS(s,    'indexes.rds')
  saveRDS(praw, 'praw.rds')
}

# Compute means across repetitions
rn <- function(x) round(x, 3)
rn(apply(s, 1:2, mean))
```

The average overfitting-corrected Brier score estimate closely tracks
the average of all the true Brier scores. The optimism bootstrap
overestimated the performance on the $D_{xy}$ and calibration slope
scales. The positive bias is 0.02 for $D_{xy}$ and 0.04 for the
calibration slope.[It is known that in extremely overfitted models, the
bootstrap will underestimate the amount of overfitting when compared
with 100 repeats of 10-fold cross-validation. But even in these
situations, model performance estimated by the bootstrap will convey
enough bad news.]{.aside}

Next examine tail non-coverage probabilities for a variety of ways of
utilizing the training and test indexes over the 300 resamples drawn for
each simulated dataset. This is done separately for each of the 3
performance indexes.

```{r}
# Add true indexes to praw
raw <- praw
for(i in 1 : nsim) raw[[i]]$true <- s[, 'true', i]

# Function applied to each element w of raw, for bootstrap type and basis and
# index number idx
# Result is a pair of logical values left and right with TRUE indicating that
# the true index is outside that tail

h <- function(w, what=c('error', 'cltrue')) {
  what  <- match.arg(what)
  true  <- w$true [idx]
  train <- w$btrain[, idx]
  test  <- w$btest [, idx]
  opt   <- train - test
  mopt  <- mean(opt)
  theta <- w$orig[idx]
  x <- switch(basis,
              train = train,
              test  = test,
              opt   = opt,
              wtd   = 2    * train - test,
              wtd2  = 1.5  * train - test,
              wtd3  = train - 0.75 * test,
              wtd4  = train - 1.25 * test,
              wtd5  = train - 1.5 * test)

  if(type == 'shifted nonpar') {
    qu <- quantile(x, c(0.025, 0.975), na.rm=TRUE)
    r <- c(qu[1] - mopt, qu[2] - mopt)
  } else if(substr(type, 1, 2) == 'sd') {
    if(basis == 'v2') {
      v1 <- var(train)
      v2 <- var(train - test)
      s  <- sqrt(v1 + v2 - 0.5 * sqrt(v1 * v2))
      r <- c(theta - mopt - s * z, theta - mopt + s * z)
    }  else {   # dualSD is in Hmisc 5.2-4
      s <- dualSD(x, nmin=if(type == 'sd1') 100000 else 10)
      a <- if(type == 'sd2') 1:2 else 2:1
      r <- c(theta - mopt - s[a[1]] * z, theta - mopt + s[a[2]] * z)
    }
  }
  if(what == 'cltrue') return(c(left=r[1], right=r[2], true=true))
  r <- c(r[1] > true, r[2] < true)
  names(r) <- .q(left, right)
  r
}

z  <- qnorm(0.975)

# validate used wtd4 sd2rev; check against that for Brier score
if(FALSE) {
  idx  <- 3
  type <- 'sd2rev'
  basis <- 'wtd4'
  for(i in c(1, nsim / 2, nsim)) {
    cat('i=', i, '\n')
    print(h(raw[[i]], 'cltrue'))
    print(s[3,,i])
  }
}

types <- c('shifted nonpar', 'sd1', 'sd2', 'sd2rev')
bases <- c('train', 'test', 'opt', 'wtd', 'wtd2', 'wtd3', 'wtd4', 'wtd5', 'v2')

D <- vector('list', 3)
for(idx in 1 :3) {
  idxname <- c('Dxy', 'Slope', 'B')[idx]
  cat('\n-------------------------------------------------------------------\nIndex: ',
      if(idxname == 'B') 'Brier' else idxname, '\n\n', sep='')

  # Analyze simulation results separately for each confidence interval construction method
  # For each simulation compute the components of confidence intervals
  # Then compute the per-simulation limits
  # Then compute for that simulation whether a non-coverage has occurred per tail

  d <- NULL
  for(type in types) {
    for(basis in bases) {
      if(type == 'shifted nonpar' && basis %nin% c('train', 'test')) next
      if(type != 'sd1'            && basis == 'v2') next
      w    <- sapply(raw, h)
      err  <- apply(w, 1, mean, na.rm=TRUE)
      dist <- sum(abs(err - 0.025))
      d <- rbind(d, data.frame(type, basis, left=rn(err[1]), right=rn(err[2]),
                               coverage=rn(1 - sum(err)), dist=rn(dist)))
    }
  }
  row.names(d) <- NULL
  print(d)
  D[[idx]] <- d
}
```

For each bootstrap confidence interval method and each performance
index, compute the sum of two distances from the target tail
probabilities of 0.025, and then sum these distances over the three
indexes and see which method has the best overall confidence interval
coverage accuracy.

```{r}
# Sum distances over the 3 indexes
di <- D[[1]]$dist + D[[2]]$dist + D[[3]]$dist
cbind(distance=di, rank=rank(di))
```

The winner (rank of 1) is `sd2rev wtd4` which computes standard
deviations of the 300 bootstrap re-sampled values of training - 1.25
$\times$ test indexes, where training denotes apparent performance on
the bootstrap sample and test denotes the bootstrap fit's performance on
the original whole sample. The simple SD of the 300 estimates of bias
(optimism; `sd1* opt`) had overall ranks of 14, 16, and 15. It stands to
reason that the best quantity on which to compute SDs puts more weight
on the test sample indexes than it does on the training sample, because
there is less variation in performance of bootstrap fits on the original
sample over simulated datasets.

`sd2rev` refers to using the "top SD" for the lower confidence limit and
the "bottom SD" for the upper confidence limit. The distance rank for
using the ordinary SD was 5.5 and the rank when reversing the two
directional SDs was 13. The standard parametric bootstrap using
approximate normality does not handle asymmetric sampling distributions,
but bottom and top SDs do. Better performance using two SDs indicates
asymmetry of sampling distributions for the indexes.

Coverage for calibration slope and Brier score was estimated to be 0.955
and 0.947 for the best overall performing method. It did not perform so
well for $D_{xy}$, giving only 0.852 coverage. $D_{xy}$ has a somewhat
degenerate distribution when computed on regression fits, which
typically do not allow $D_{xy} < 0$, so normality is usually not a good
fit for $D_{xy}$.

# Accuracy of Estimated Calibration Curves and Confidence Limits

Let's study how well the overall winning bootstrap confidence interval
estimator for three regular performance indexes fares when computing
confidence bands for a vector of parameters, bootstrap
overfitting-corrected calibration curves. Let's fix the grid of 50
equally spaced points for which the calibration curve is estimated,
using the 0.02 and 0.98 quantiles of pooled distributions of predicted
probabilities for 10 repeated model fits. The simulations we are about
to run are similar to those for the 3 statistical indexes, but now we
have 50 indexes.

A histogram of the pooled distribution shows that the entire range of
predicted probabilities is represented by the chosen logistic model and
distribution of $x_1$.

```{r}
set.seed(1)
phat <- replicate(10, {
  d <- simd(N, np)
  f <- lrm(y ~ x, data=d)
  predict(f, type='fitted')
})
round(apply(phat, 2, quantile, c(0, 0.02, 0.98, 1)), 2)
hist(phat, nclass=50, main=NULL, xlab=expression(hat(P)))
# On average for n=200, request calibrated probabilities for 5th to 196th
# largest observed predicted risks, which is pushing things a bit
r <- quantile(phat, probs=c(0.02, 0.98))
cat('Limits on predicted probabilities for which calibration curve is estimated:',
    round(r, 3), '\n')
pp <- seq(r[1], r[2], length=50)
```

For a given logistic model fit `f` on dataset `d` compute the true
calibration curve at `pp`. The calibration curve is a set of points
$(\hat{P}, P)$, where $\hat{P}$ are predicted probabilities and $P$ are
true probabilities that $Y=1$ given $\hat{P}$. The points are fitted
using least squares in the logit (the true calibration is linear in the
logits by design) to average over variation in
$X_{2}, \ldots, X_{`r np`}$ that yields unnecessary variation in
$\hat{P}$ that arises from giving the last $p - 1$ (noise) $X$'s a
chance to get nonzero $\hat{\beta}$'s. So the true calibration curve is
an average over covariate distributions, conditioning only on the
covariates through conditioning on $\hat{P}$.

```{r}
truecal <- function(f, d, pl=FALSE) {
  Lhat  <- predict(f)
  Ltrue <- d$x[, 1]
  phat  <- plogis(Lhat)
  int_slope <- lm.fit.qr.bare(Lhat, Ltrue)$coefficients
  ptrue <- plogis(Ltrue)
  if(pl) plot(phat, ptrue, col='gray80')
  plogis(int_slope[1] + int_slope[2] * qlogis(pp))
}

# Test the function
par(mfrow=c(3,3), mar=c(2,2,0,0))
set.seed(3)
g <- function() {
  d <- simd(N, np)
  f <- lrm(y ~ x, data=d)
  tcal <- truecal(f, d, pl=TRUE)
  lines(pp, tcal)
  abline(a=0, b=1, col='gray80')
  coef(f)
}
round(replicate(9, g()), 2)
```

The fact that the circles are sometimes far from the curves implies that
an "average" true calibration curve is hiding some false variation in
predicted risks due to variation in the extraneous predictors. In other
words, many different predicted values (depending on the noise
predictors) can give rise to one true probability. To have a gold
standard to compare with bootstrap estimates we must take averages.

The default calibration estimator used in `calibrate()` is *loess*,
using the R `lowess` function with outlier detection turned off
(`iter=0`).

Define alternative parametric linear and quadratic (in the logit)
estimators.

```{r}
qcal <- function(x, y, deg=2) {
  # Curtail predicted probabilities to [0.001, 0.999]
  x <- pmax(0.001, pmin(x, 0.999))
  L <- qlogis(x)
  f <- lrm.fit(if(deg == 2) cbind(L, L^2) else L, y, compstats=FALSE)
  k <- coef(f)
  pgrid <- seq(min(x), max(x), length=200)
  Lgrid <- qlogis(pgrid)
  cgrid <- k[1] + k[2] * Lgrid
  if(deg == 2) cgrid <- cgrid + k[3] * Lgrid^2
  cal   <- plogis(cgrid)
  list(x=pgrid, y=cal)
}
```

Test the linear and quadratic calibrations in a setting where the
calibration curve is known to be the line of identity, and also include
*loess* estimates (red). Linear logistic calibration uses a black line,
and quadratic a blue line.

```{r}
par(mfrow=c(3, 3), mar=c(2, 2, 0, 0))
set.seed(1)
for(i in 1 : 9) {
  x <- runif(N)
  y <- ifelse(runif(N) <= x, 1, 0)
  w1 <- qcal(x, y, deg=1)
  plot(w1, type='l', xlim=c(0, 1), ylim=c(0, 1))
  w2 <- qcal(x, y, deg=2)
  lines(w2, col='blue')
  l <- lowess(x, y, iter=0)
  lines(l, col='red')
  abline(a=0, b=1, col='grey80')
}
```

Define a function that draws one sample, fits a binary logistic model,
and computes bootstrap percentile confidence bands for the calibration
curve at a regular sequence of predicted probabilities `pp`. Add the
true calibration curve to the output, alongside $\min(a, b)$, where $a$
is number of observed predicted risks that are less than the current
value of `pp` and $b$ is the number of $\hat{P}$ that are greater than
the value of `pp`. In addition, save the standard deviation of
$\hat{P}$. This is done for one calibration estimator. A special case is
`smoother='basic'` which stands for using the `validate` function to get
bias-corrected intercept and slope estimates and to convert them to
linear (in the logit) calibration curves, without confidence limits.
Finally, `smoother='apparent'` does not use bootstrapping and instead
takes the $\hat{P}$ to (falsely) need no calibration, to gauge the
accuracy of uncalibrated estimates.

```{r}
# Use the following if you want to count "close" observations
# pptab <- data.table(pp, low = pp - 0.05, hi = pp + 0.05)
pptab <- data.table(ppgrid = pp)
sim1 <- function(n, p, B=300, smoother='lowess', pl=FALSE) {
  .i. <<- .i. + 1
  cat(.i., '', file='/tmp/z', append=TRUE)
  d <- simd(n, p)
  f <- lrm(y ~ x, data=d, x=TRUE, y=TRUE)
  phat    <- predict(f, type='fitted')
  ptrue   <- truecal(f, d)
  if(is.character(smoother) && smoother != 'lowess') {
    switch(smoother,
           basic = {
            val       <- validate(f, B=B)[, 'index.corrected']
            int_slope <- val[c('Intercept', 'Slope')]
            estcal    <- plogis(int_slope[1] + int_slope[2] * qlogis(pp)) },
           apparent = {
             estcal   <- pp } )
    lower <- NA
    upper <- NA
    if(pl) {
      plot(pp, estcal, type='l', xlim=c(0,1), ylim=c(0,1))
      lines(pp, ptrue, col='blue')
    }
  } else {
    bootcal <- calibrate(f, B=B, predy=pp, smoother=smoother)
    estcal <- bootcal[, 'calibrated.corrected']
    lower  <- pp + bootcal[, 'Lower']
    upper  <- pp + bootcal[, 'Upper']
    if(pl) {
      plot(bootcal, ylim=c(0,1))
      lines(pp, ptrue,   col='blue')
    }
  }
  d      <- data.table(phat)
  lo <- d[pptab, on = .(phat <= ppgrid), .N, by=.EACHI][, N]
  hi <- d[pptab, on = .(phat >= ppgrid), .N, by=.EACHI][, N]
  # count  <- d[pptab, on = .(phat >= low, phat <= hi), .N, by = .EACHI][, N]
  invisible(cbind(pp, estcal, ptrue, lower, upper, count=pmin(lo, hi), sphat=sd(phat)))
}
```

Run 9 tests for the default *loess* calibration estimator, plotting the
confidence bands. The blue curve is the true calibration curve. In the
upper left of each panel show the number of observations with low and
high predicted probabilities.

```{r}
#| column: screen-inset
#| fig-height: 8
#| fig-width: 8
set.seed(9)
par(mfrow=c(3,3), mar=c(2, 2, 0, 0))
for(j in 1 : 9) sim1(N, np, pl=TRUE)
```

Run 1000 simulations using all three calibration curve estimators. For
each dataset simulated, record the number of observations with low/high
predictions. Also compute the more direct estimated bias-corrected
linear calibration curve by using the `validate` function to debias the
intercept and slope computed from bootstrapping that is separate from
the `calibrate` run.

```{r}
sm <- list('lowess',
           function(x, y) qcal(x, y, deg=1),
           function(x, y) qcal(x, y, deg=2),
           'basic',
           'apparent')
set.seed(7)
if(file.exists('cal.rds')) R <- readRDS('cal.rds') else {
  R        <- vector('list', 5)
  names(R) <- c('loess', 'linear', 'quadratic', 'basic', 'apparent')
  for(est in 1 : 5) {
    cat('\nEstimator', est, '\n')
    .i. <- 0
    r <- replicate(1000, sim1(N, np, B=300, smoother=sm[[est]]))  # 44m
    R[[est]] <- r
  }
  saveRDS(R, 'cal.rds')
}
```

Restructure the results into a list of data tables.

```{r}
convert2dt <- function(r) {
  npp  <- length(pp)
  nsim <- dim(r)[3]
  vn   <- dimnames(r)[[2]]
  # Restructure array so that pp and simulation # move fastest
  r <- aperm(r, c(1, 3, 2))
  # Stack into a taller matrix
  r <- matrix(r, nrow = npp * nsim, ncol=length(vn))
  d <- data.table(pp  = rep(pp, times=nsim),
                  sim = rep(1:nsim, each=npp) )
  d[, (vn) := as.data.table(r)]
  d
}
d <- lapply(R, convert2dt)
```

## Accuracy of Bootstrap Calibration Estimates for $n=200$

Group the number of points within 0.05 of the target estimated risk
$\hat{P}$ into intervals and compute pseudo-medians[The pseudo-median is
the median of all possible pairwise means. It is more robust than the
mean and is virtually as efficient as the mean, unlike the sample
median.]{.aside} of absolute calibration curve estimation error by
calibration estimator, target $\hat{P}$, and group, where groups are
intervals of $\hat{P}$ SDs or minimum tail frequencies.

```{r}
pacc <- function(d, stratby) {
  # Restructure to a single tall and thin data table
  w <- rbindlist(d, idcol='estimator')
  if(! missing(stratby)) {
    w[, let(z         = abs(estcal - ptrue),
            group     = eval(stratby) ) ]
    w[, prn(table(group))]
    wall <- copy(w)
    wall[, group := 'All']
    w <- rbind(w, wall)
    w <- w[, .(err = pMedian(z, na.rm=TRUE)), by=.(estimator, group, pp)]
    g <- ggplot(w, aes(x=pp, y=err, color=estimator)) +
      facet_wrap(~ group)
  } else{
    w[, let(z = abs(estcal - ptrue))]
    w <- w[, .(err = pMedian(z, na.rm=TRUE)), by=.(estimator, pp)]
    g <- ggplot(w, aes(x=pp, y=err, color=estimator))
  }
  g + geom_line() +
    xlab(expression(hat(P))) + ylab('pMedian |Calibration Estimation Error|')
}
pacc(d, expression(cut2(sphat, c(.175, .2, .25)))) + labs(caption='By groups of SD(Phat), n=200')
pacc(d, expression(cut2(count, 50))) + labs(caption='By groups of min(tail count), n=200')
```

It is typically true that even in heavily overfitted models, predictions
that are near the grand mean estimated risk are accurate. This is
reflected in the above plots which show the "no calibration" method of
just trusting the apparent predicted risks to have the lowest expected
error in the middle of the risk distribution. But it is highly
overoptimistic for outer risks.

Run the following if you want to see a tensor spline surface for
$x = \hat{P}$, $y=$ SD of $\hat{P}$.

```{r eval=FALSE}
require(mgcv)
for(nam in names(d)) {
  w <- d[[nam]]
  f <- gam(abs(estcal - ptrue) ~ te(pp, sphat), data=w)
  plot(f, scheme=2, main=nam)
}
```

## Accuracy of Approximate Bootstrap Confidence Limits for Calibration, $n=200$

For each value of `pp` compute the proportion of simulations for which
`lower > ptrue` and the proportion `upper < ptrue` (tail non-coverage
proportions). Also compute overall error proportions.

```{r}
plotit <- function(R, title) {
  left  <- R[, 'lower', ] > R[, 'ptrue', ]
  right <- R[, 'upper', ] < R[, 'ptrue', ]

  rmn <- function(x) round(mean(x, na.rm=TRUE), 3)
  cat('Overall non-coverage for', title, ':\tleft=', rmn(left),
      ' right=', rmn(right), '\n')

  left  <- apply(left , 1, mean, na.rm=TRUE)
  right <- apply(right, 1, mean, na.rm=TRUE)
  nc    <- c(left, right)
  phat  <- c(pp, pp)
  k     <- length(left)
  side  <- c(rep('left', k), rep('right', k))
  ggplot(mapping=aes(x=phat, y=nc, col=side)) + geom_line() + xlab(expression(hat(P))) +
      ylab('Non-Coverage') + ylim(0, 0.1) + labs(caption=title)
  }
for(nam in .q(loess, linear, quadratic))
  print(plotit(R[[nam]], title=nam))
```

The desired non-coverage probability is 0.025, which is achieved for
interior predicted probabilities. The problem with coverage
probabilities in the tails seems to be due to wild estimates (especially
for *loess*) or extrapolations at more rare predicted probabilities,
i.e., those closer to 0 or 1. Linear logistic calibration worked better
than the more flexible calibration estimators. Note that the simulations
were stacked in favor of linear calibration.

## Calibration for $n=800$ and $n=1500$

Try two larger sample sizes.

```{r}
set.seed(8)
Ns <- c(800, 1500)
for(N in Ns) {
  fi <- paste0('cal', N, '.rds')
  if(file.exists(fi)) Rn <- readRDS(fi) else {
    Rn        <- vector('list', 5)
    names(Rn) <- c('loess', 'linear', 'quadratic', 'basic', 'apparent')
    for(est in 1 : 5) {
      cat('\nEstimator', est, '\n')
      i <- 0
      r <- replicate(1000, sim1(N, np, B=300, smoother=sm[[est]]))  # 91m
      Rn[[est]] <- r
    }
    saveRDS(Rn, fi)
  }
  if(N ==  800) R800  <- Rn
  if(N == 1500) R1500 <- Rn
}
```

Again compute accuracies of bias-corrected calibration estimates.

```{r}
for(N in Ns) {
  Rn <- if(N == 800) R800 else R1500
  d <- lapply(Rn, convert2dt)
  cap <- paste0('By groups of SD(Phat), n=', N)
  g <- pacc(d, stratby=expression(cut2(sphat, g=4))) + labs(caption=cap)
  print(g)
  cap <- paste0('By groups of min(tail counts), n=', N)
  g <- pacc(d, stratby=expression(cut2(count, c(20, 50, 100)))) + labs(caption=cap)
  print(g)
}
```

Now display confidence interval non-coverage proportions.

```{r}
for(N in Ns) {
  Rn <- if(N == 800) R800 else R1500
  for(nam in .q(loess, linear, quadratic))
    print(plotit(Rn[[nam]], title=paste0(nam, ', N=', N)))
}
```

Especially for linear and quadratic calibration, increasing the sample
size made coverage larger than 0.95 for linear and quadratic
calibration. Perhaps it's best to have this conservatism so that the
method will not be anti-conservative for small sample sizes.

# Summary

I considered approximate confidence bands for the true model performance
indexes and entire calibration curves as estimated by a bootstrap
overfitting-corrected approach. The dual standard deviation approximate
bootstrap confidence interval ABCLOC balances tail non-coverage
probabilities by allowing the bootstrap distribution of estimated biases
to be asymmetric. ABCLOC seems to provide accurate enough confidence
limits for individual accuracy scores such as the Brier score,
calibration slope, and to a lesser extent Somers' $D_{xy}$, and I have a
similar expectation of good performance for other scaler indexes.
Considering vector measures such as whole calibration curves evaluated
over a sequence of predicted risks, ABCLOC can be conservative. ABCLOC
intervals are useful, and are certainly better than having no
uncertainty intervals. They require no additional computation time.

I hope that others will refine ABCLOC, and run more diverse simulations.

The choice of calibration curve estimator also matters, and an important
point demonstrated by simulation is the difficulty of estimating
calibration (not to mention its uncertainty) at very low or high
predicted probabilities. The more flexible *loess* method has worse
confidence interval coverage, especially for predicted risks far from
the mean. When the true calibration curve is logit-linear, linear
calibration estimates, fitted by running binary logistic regression with
a sole predictor equal to the logit of predicted risk, are superior as
expected. Quadratic calibration is a good compromise, and the
performance of restricted cubic spline estimators with 4 knots should
also be explored. An advantage of parametric calibration methods
(linear, quadratic, spline, ...) is that just the parameters (3 for
quadratic) of the estimated calibration curve can be saved for each
bootstrap fit, and one does not need to track the individual calibration
curve points (50 in number in simulations presented here) as must be
done for *loess*.[At present, the approach of saving only the
calibration parameters is only implemented in `validate()` functions for
calibration intercept and slope.]{.aside}

# Computing Environment

```{r}
sessionInfo()
```
